---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

`sraplus` is a flexible assessment package based around Ovando et al. 2019. At the most "data limited" end, the model approximates the behavior of catch-msy, sampling from prior distributions to obtain parameter values that given a catch history do not crash the population and are match supplied priors on initial and final depletion. At the most data rich end the model the model can fit to an abundance index with priors on fishing mortality over time and initial and final depletion.


## Installing

This is an in-development package hosted on github, so you will need to do a few things to install it. 

1. open R

2. If you don't have the `devtools` package installed yet, run

```{r, eval = FALSE}
install.packages("devtools")
```

You'll need to be connected to the internet.

3. Once `devtools` has been installed, you can then install `sraplus` by running

```{r, eval = FALSE}
devtools::install_github("danovando/sraplus")
```

That's probably going to ask you to install many other packages, agree to the prompts. 

## SIR example

Once you've succesfully installed `sraplus` you can take for a test drive with these examples. 

For the first example we'll run use a sampling-importance-resampling (SIR) algorithm, using fisheries management index scores and swept area ratio data to provide priors on B/Bmsy and U/Umsy in the final year

```{r}

library(sraplus)
library(ggplot2)
library(dplyr)

example_taxa <- "gadus morhua"

data(cod)

driors <- format_driors(
  taxa = example_taxa,
  catch = cod$catch,
  years = cod$year,
  initial_b = 1,
  terminal_b = 0.5,
  use_heuristics = TRUE,
  sar = 2,
  fmi = c("research" = 0.5,"management" = 0.8, "enforcement" = 0.8, "socioeconomics" = 0.9)
)

plot_driors(driors)

sir_fit <- fit_sraplus(driors = driors,
                       use_sir = TRUE,
                       draws = 1e4)

plot_sraplus(sir_fit = sir_fit)

```


## Maximum Likelihood Example

We'll now use maximum likelihood to fit to the same fishery, but now using an index of abundance and a swept area ratio as a penalty on U/Umsy in the final year

```{r}

driors <- format_driors(taxa = example_taxa,
                                 catch = cod$catch,
                                 years = cod$year,
                        index = cod$index,
                        index_years = cod$year,
                        initial_b = 1,
                        initial_b_sd = 0.01,
                        terminal_b = 0.5,
                        sar = 2,
                        sar_sd = 0.1)

plot_driors(driors)


ml_fit <- fit_sraplus(driors = driors,
                       use_sir = FALSE,
                       model = "sraplus_tmb")
```


And we can now compare the two results


```{r}

plot_sraplus(sir_fit = sir_fit, ml_fit = ml_fit)

```

